---
author: "Joseph Diaz"
title: "Lab 5: Simulating discrete probability models"
date: "Spring 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



### The problem
The code chunk above performs the experiment once: spin the roulette wheel until you lose and record the number of spins.  Simulate 1000 experiments.  As usual, do this by wrapping the code chunk above within a for-loop and storing the number of spins `streak` in a vector.
```{r rand, echo=TRUE}
oneSpin = function(p)
{
  streak = 0
  while(TRUE)
  {
    streak = streak + 1 # tally streak
    if(runif(1) < p) # generate Bernoulli with probability p
      break 
  }
  return(streak)
}

streakers = function(n,p)
{
  streaks = replicate(n, oneSpin(p))
  return(streaks)
}
p = 20/38
n = 1000
```

### Report the following:

* Histogram of the win streak length.  Note that this is a discrete distribution so should place histogram bars at discrete values {0, 1, 2, ...}.  This may be done with the `breaks` option within `hist`.  If your storage variable is called `winstreak`:
```{r hist, echo=TRUE}
winstreaks = streakers(n, p)
hist(winstreaks, br=seq(min(winstreaks)-0.5, max(winstreaks)+0.5))
```

* Average length of the win streak: *`r mean(winstreaks)`*

* Standard deviation of the winning streak lengths: *`r sqrt(var(winstreaks))`*

* Compare the empirical average and standard deviation in the previous two bullets to the true values from the Geometric(*p* = 20/38) distribution.  
  *The theoretical expected value and standard deviation are*
$$E[X] = \frac{1}{p} \approx `r 1/p`,\quad SD[X] = \sqrt{\frac{1-p}{p^2}} \approx `r sqrt((1-p)/(p^2))`$$
    *Which are reasonably close to the empirical mean and standard deviation we calculated.*

* Longest winning streak. *`r max(winstreaks)`*

\vfill

## Task 2: Simulating negative binomial distributions

In this task, we will compare two different algorithms for simulating from a negative binomial distribution.

### Problem (a)
Recall that a negative binomial random variable *NB(r, p)* is the sum of *r Geometric(p)* random variables.  Use the algorithm from Task 1 to simulate 1000 *NB(10, 0.6)* random variates.


```{r codecore2, echo=TRUE}
sumOfGeo= function(r, p)
{
  runs = 0
  trials = 0
  while(runs < r)
  {
    trials = trials + oneSpin(p)
    runs = runs + 1
  }
  return(trials)
}
n = 10000
p = .6
r = 10

```

### Problem (b)
The negative binomial pmf induces the following recursion relation.  If $X \sim NB(r,\ p)$, then
\[ P(X = i+1) = \frac{i(1-p)}{(i+1-r)} \cdot P(X=i),\quad i \geq r\]
Use this recursion relation to generate 1000 $NB(10,\ 0.6)$ random variates.


```{r NegBin, echo=TRUE}
rTrials = function(p,r)
{
  pmf = p^r
  cdf = pmf
  j = r
  u = runif(1)
  while(u >= cdf)
  {
    pmf = ((j*(1-p))/(j+1-r))*pmf
    cdf = cdf + pmf
    j = j + 1
  }
  return(j)
}
NegBinSim = function(p, n, r)
{
  return(replicate(n, rTrials(p,r)))
}

trials = NegBinSim(p, n, r)
sumGeo = replicate(n, sumOfGeo(r, p))
```


This binomial simulator may be applied directly after changing just three lines: 

* $j=r$

* the recursion relation formula

* `pmf = p^r`

### Report the following for each of the simulations in problems (a) and (b)

* Histogram of the variates
```{r histy, echo=TRUE}
hist(trials, br=seq(min(trials)-0.5, max(trials)+0.5))
hist(sumGeo, br=seq(min(sumGeo)-0.5, max(sumGeo)+0.5))
```
 * Mean and standard deviation of the simulated variates  
  *Mean from NegBin: `r mean(trials)`, Mean from sumGeo: `r mean(sumGeo)`*  
  *SD from NegBin: `r sqrt(var(trials))`, SD from sumGeo: `r sqrt(var(sumGeo))`*

* Run time: compare computing speed between the two algorithms.  In R, can wrap your algorithm or sequence of operations as follows to time your code.
```{r timewrapper}
x = proc.time()
sumGeo = replicate(n, sumOfGeo(r, p))
timer = proc.time() - x
algtimeGeo = timer[3]

x = proc.time()
trials = NegBinSim(p, n, r)
timer = proc.time() - x
algtimeNegBin = timer[3]
```

### Questions

* How do the histograms compare?

*They are practically identical.*

* How do the mean and standard deviation from the simulations compare to the true mean and standard deviation of a $NB(0.6,\ 10)$ distribution?

*The theoretical mean and standard deviation are*
$$E[X] = `r round(r/p, digits = 3)`, \quad SD[X] = `r round(sqrt(r*(1-p)/p^2), digits = 3)`$$
  *which is very close to the simulated mean and standard deviation displayed under the histograms.*

* How do the computing times compare?  Which algorithm is faster?

*sumGeo time: `r algtimeGeo`, NegBin time: `r algtimeNegBin`. The Negative Binomial method was much lower; it's much faster.*

* "Simulation flops": Which simulator do you think uses more uniform random numbers (call to the `runif()` function)?  Why?

*The Geometric simulation uses a lot more calls of `runif()`, as a call is made for each individual geometric trial; as opposed to Negative Binomial method which uses a call to `runif()` for what is the equivalent to 10 geometric trials.*