---
author: "Joseph Diaz"
title: "Final: Coding Portion"
date: "Spring 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



### Task 1:  Variation on the coupon collector’s problem
On Opening Day of baseball, we will look at the top 18 players on each of the 30 major league teams. I want to collect baseball cards for each of these players, so, there are $n = 540$ baseball cards in all.  I can buy a pack of baseball cards, which contain five distinct players in each pack.  We want to know how many packs a person needs to buy in order to complete the collection, presuming no trading of cards or other strategies to obtain cards missing from a collection. 

* Simulate this baseball card collection problem 1000 times. Keep track of the number of packs required to complete the collection of 540 baseball cards.
```{r cardgame}
collectAll = function()
{
  k = 0
  collected = vector()
  while(!all(1:540 %in% collected))
  {
    collected = c(collected, sample(1:540, 5))
    k = k + 1
  }
  return(k)
}
packs = replicate(1000, collectAll())
```

* Report the smallest and largest number of packs required as well as the mean and standard deviation of the number of packs from your simulation.  
$$\begin{tabular}{|c|c|c|c|} \hline Max  & Min & $\mu$ & $\sigma$  \\
                        \hline `r max(packs)` & `r min(packs)` & `r round(mean(packs), digits=3)` & `r round(sqrt(var(packs)), digits = 3)` \\
                        \hline\end{tabular}$$

* Present a histogram of the simulated number of packs.
```{r histogram}
hist(packs, breaks=seq(min(packs),max(packs),(max(packs)-min(packs))/(50)))
```

* Present a plot of the complimentary cumulative distribution function of your simulation in (a), that is, a plot of the empirical probability the number of packs collected exceeds values $t > 0$ (this is the complement of the cdf). (Hint: use the r function ecdf()).  
```{r ecdf}
fn = ecdf(packs)
plot(sort(1-fn(1:1000), decreasing = TRUE), type = "l", col='blue')
title("CCDF of packs")
dobrow = function(n){return((n/5)*sum(1/1:n))}
```

* Use the approximation from Dobrow Section 5.1.2, deriving the expected number of packs needed to complete the baseball card collection from a sum of geometric random variables:
$$E(\text{\# of packages}) \approx \frac{n}{5}\cdot\left(1 + \frac{1}{2} + \frac{1}{3} + \cdots + \frac{1}{n}\right) \approx \frac{n\ln n}{5}$$
Compare this approximation to your Monte Carlo estimate in (b). Which do you trust more and why?  
  
The approximation gives: $E[X] \approx `r dobrow(540)`$. Which is very close to the empirical mean of `r mean(packs)` calculated above; the percent error between them only being about `r (abs(dobrow(540) - mean(packs))/dobrow(540))*100`%. I trust the empirical mean more, because its the consquence of actual measurements from actual experiments. Theoretical approximations have their uses, but hard data is more convincing.

### Task 2:
Let $S \sim \text{Uniform}(10, 14)$, Let $A$ be the area of a regular hexagon with side length, $S$. Use R to simulate $S$. Simulate the mean and pdf of $A$ and compare to the exact results. Create one graph with both the theoretical density and the simulated distribution.   
  
We have that the area of a regular hexagon with side lengths $S$ is $A = \frac{3\sqrt{3}}{2}\cdot S^2$ with cdf and pdf:
$$F(x) = \frac{\sqrt{\frac{2}{3\sqrt{3}}x}-10}{4} \implies F'(x) = f(x) = \frac{1}{4\sqrt{6x\sqrt{3}}}$$
```{r HexaSim}
n = 1000000
S = runif(n, min = 10, max = 14)
A = (3/2)*sqrt(3)*S^2
meanA = mean(A)
varA = var(A)
maxA = max(A)
minA = min(A)
pdf= function(x)
{
  return(1/(4*sqrt(6*x*sqrt(3))))
}
x = seq(minA,maxA,by=((maxA-minA)/(n/10000-1)))
hist(A, breaks=x, probability = TRUE, main = "Histogram of A with theoretical PDF of A")
curve(pdf(x), add=T, lw=3, col='red')
```
We have that the empirical mean of $A$ is `r round(meanA,digits=4)`. While the theoretical mean of $A$ is  
$$E[A] = \int_{150\sqrt{3}}^{294\sqrt{3}}xf(x)\ dx =\frac{1}{4\sqrt{6\sqrt{3}}}\int_{150\sqrt{3}}^{294\sqrt{3}} \sqrt{x}\ dx = \frac{2}{3\cdot4\sqrt{6\sqrt{3}}}\left[x^{3/2}\right]_{150\sqrt{3}}^{294\sqrt{3}}$$
$$\Rightarrow = \frac{(294\sqrt{3})^{3/2}- (150\sqrt{3})^{3/2}}{6\sqrt{6\sqrt{3}}} \approx 377.5871$$
These are pretty close, with the percent error between them being only about `r (abs(377.58707 - meanA)/377.58707)*100`%.

### Task 3:
Conduct a simulation study to illustrate that sums of independent exponential random variables produce a gamma distribution. In particular, let $X_1, \cdots, X_5$ be exponentially distributed with $\lambda = 1$. Simulate $X_1 + \cdots + X_5$. Plot a histogram estimate of the distribution of the sum together with the exact density function of a gamma distributed random variable with $a = 5, \lambda = 1$.

```{r expToGamma}
G = replicate(n, sum(rexp(5,1)))
hist(G, breaks = seq(min(G), max(G), (max(G)-min(G))/(n/10000-1)), probability = TRUE, 
     main = "Histogram of G with theoretical pdf of gamma(5,1)")
curve(dgamma(x,5,1), 0, 20, add=TRUE, lw=3, col='red')
```