---
title: "Homework 4"
author: "Joseph Diaz"
date: "February 13, 2020"
output: pdf_document
---

```{r}
n = 100000
```

\subsection*{1.41}
Modify the code in the R script \textbf{CoinFlip.R} to simulate the probability of
getting exactly one head in four coin tosses.
```{r }
CoinFlip4 = function(n)
{
  flip = matrix(sample(0:1, 4*n, replace = TRUE), nrow = 4, byrow = FALSE)
  total = flip[1,1:n] + flip[2,1:n] + flip[3,1:n] + flip[4,1:n]
  mean(total == 1)
}

CoinFlip4(n)
```

\subsection*{1.42}
Modify the code in the R script \textbf{Divisible356.R} to simulate the probability
that a random integer between 1 and 5000 is divisible by 4, 7, or 10. Compare
with your answer in Exercise 1.37.
```{r}
Divisible4710 = function(n)
{
  num = sample(1:5000, n, replace = TRUE)
  mean(num%%4==0|num%%7==0|num%%10==0)
}

Divisible4710(n)
```

\subsection*{1.43}
Use R to simulate the probability of getting 8 in the sum of two
dice rolls.
```{r}
SumGreaterEqual8 = function(n)
{
  rolls = matrix(sample(1:6, 2*n, replace = TRUE), nrow = 2, byrow = FALSE)
  total = rolls[1,1:n] + rolls[2,1:n]
  mean(total == 8)
}

SumGreaterEqual8(n)
```

\subsection*{1.44}
Use R to simulate the probability at least one 2 in 5 rolls of a tetrahedron dice.
```{r}
TetraDice = function(n)
{
  rolls = matrix(sample(1:4, 5*n, replace = TRUE), nrow = 5, byrow = FALSE)
  success = numeric(n)
  for(i in 1:n)
  {
    success[i] = if(2 %in% rolls[1:5,i]) 1 else 0
  }
  mean(success)
}

TetraDice(n)
```
