---
title: 'Lab 4: Simulation'
author: 'Joseph Diaz'
output:
  pdf_document
date: "Spring 2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### The problem

In this task, we will evaluate an original implementation of the `rand` command in *Matlab* for generating uniform random numbers.  The random number generator has formulation

\[ X_{n+1} = 16807 X_n mod (2^{31}-1).\]

Ammend the code chunk above to generate 1000 random uniform variates from this random number generator.

```{r rand, echo=FALSE}

rando = function(n, seed)
{
  for(i in 1:n)
  {
    seed<-c(seed, (65539*seed[i])%%(2^31) - 1)
  }
  return(seed[2:n])
}
x = rando(1000,1)
#par(mfrow=c(2,1))
# histogram of the n RANDU variates
hist(x, main="", xlab="RANDU variates", ylab="Frequencies")

# empirical distribution function of the n RANDU variates and
# uniform(0, 1) probability plot
plot.ecdf(x, verticals= TRUE, do.p = FALSE, main="", ylab="Probability")
abline(0,1)

# Kolmogorov-Smirnov test of RANDU variates against U(0, 1) distribution
ks.test(x,"punif",0,1)
```

### Questions:

* Do the pseudo-random numbers appear uniformly distributed?  Why or why not?

*They do. The height of the bars on the histogram are approximately the same, which is indicative of each random value having a similar frequency. And the second plot mirrors a uniform probability plot.0*

* What are your conclusions from the Kolmogorov-Smirnov test at the $\alpha = 0.05$ level?

*We have that the p-value is less than 5%, so we reject the null hypothesis.*

## Task 2: Simulation of discrete distributions

A bag contains one red, two blue, three green, and four yellow marbles. A sample of three marbles is taken without replacement. Let B denote the number of blue marbles and Y denote the number of yellow marbles in the sample. The probabilities of each outcome is as follows:

### The problem

Perform a simulation experiment of drawing three marbles from the bag 10,000 times.  I have initialized storage vectors for you.
```{r task2, echo=TRUE}
# True values:
# P(0 blue) = 56/120; P(1 blue) = 56/120; P(2 blue) = 8/120
# P(0 Y) = 20/120; P(1 Y) = 60/120; P(2 Y) = 36/120; P(3 Y) = 4/120
marbs = function()
{
  samp = 
    sample(c("r","b","b","g","g","g","y","y","y","y"),3)
  blues = sum(samp=="b") 
  yellows = sum(samp=="y") 
  return(c(blues, yellows))
}

simnum = 10000 
bluyel = replicate(simnum, marbs())
blue = bluyel[1,1:simnum]
yellow = bluyel[2, 1:simnum]


t = table(blue)
b = append(t, NA, after=length(blue))
s = table(yellow)

disp = data.frame(b,s)
names(disp) = c("Blue Marbles","  ", "Yellow Marbles")
disp
```

### Reporting and questions

By empirical pmf (probability mass function), we mean the proportion of experiments in which we draw 0 marbles of the given color, 1 marble of the given color, 2 marbles of the given color, and 3 marbles of the given color.  So if $B$ is a random variable denoting the number of blue marbles drawn, the empirical pmf is an estimate of $P(B=0)$, $P(B=1)$, $P(B=2)$, and $P(B=3)$.

* Present the empirical pmf of the number of blue marbles drawn.
$$P(B = 0) = \frac{`r sum(blue == 0)`}{10000} = `r round(sum(blue == 0)/100, digits = 3)`\%$$
$$P(B = 1) = \frac{`r sum(blue == 1)`}{10000} = `r round(sum(blue == 1)/100, digits = 3)`\%$$
$$P(B = 2) = \frac{`r sum(blue == 2)`}{10000} = `r round(sum(blue == 2)/100, digits = 3)`\%$$

* Present the empirical pmf of the number of yellow marbles drawn.
$$P(Y = 0) = \frac{`r sum(yellow == 0)`}{10000} = `r round(sum(yellow == 0)/100, digits = 3)`\%$$
$$P(Y = 1) = \frac{`r sum(yellow == 1)`}{10000} = `r round(sum(yellow == 1)/100, digits = 3)`\%$$
$$P(Y = 2) = \frac{`r sum(yellow == 2)`}{10000} = `r round(sum(yellow == 2)/100, digits = 3)`\%$$
$$P(Y = 3) = \frac{`r sum(yellow == 3)`}{10000} = `r round(sum(yellow == 3)/100, digits = 3)`\%$$

* Present the mean and variance of the number of blue marbles drawn.
$$E[B] = `r mean(blue)`,\quad V[B] = `r var(blue)`$$

* Present the mean and variance of the number of yellow marbles drawn.
$$E[Y] = `r mean(yellow)`,\quad V[Y] = `r var(yellow)`$$

* Compare the empirical pmf with the true values.

\begin{align*}
\text{Empirical} \quad  & \quad \text{True}\\
P(B = 0) = `r round(sum(blue == 0)/10000, digits = 3)`,\quad & \quad P(B = 0) = \frac{56}{120} = `r 56/120`\\
P(B = 1) = `r round(sum(blue == 1)/10000, digits = 3)`,\quad & \quad P(B = 1) = \frac{56}{120} = `r 56/120`\\
P(B = 2) = `r round(sum(blue == 2)/10000, digits = 3)`,\quad & \quad P(B = 2) = \frac{8}{120} = `r 8/120`\\
P(Y = 0) = `r round(sum(yellow == 0)/10000, digits = 3)`,\quad & \quad P(Y = 0) = \frac{20}{120} = `r 20/120`\\
P(Y = 1) = `r round(sum(yellow == 1)/10000, digits = 3)`,\quad & \quad P(Y = 1) = \frac{60}{120} = `r 60/120`\\
P(Y = 2) = `r round(sum(yellow == 2)/10000, digits = 3)`,\quad & \quad P(Y = 2) = \frac{36}{120} = `r 36/120`\\
P(Y = 3) = `r round(sum(yellow == 3)/10000, digits = 3)`,\quad & \quad P(Y = 3) = \frac{4}{120} = `r 4/120`\\
\end{align*}

