---
title: 'Lab 3: Conditional Probability'
author: 'Joseph Diaz'
output:
  pdf_document
date: "Spring 2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## On for-loops and if-else statements for probability estimates in simulation experiments
In the last lab, we used a few slick R functions to simulate experiments and compute empirical probabilities.  But we often have to resort to, what I call, "brute-force" for-loops to repeat a simulation experiment and then Boolean expressions or if-else statements to identify successful events in experiments and estimate probabilities.  In fact, we did this in Task 5 of the "Introduction to R" lab practical.

### Code set-up
Consider the simple experiment of rolling a six-sided die and estimating the probability of rolling a 2.  Of course we know the answer is 1/6 and we can simulate this experiment in R without for-loops.  But as an illustration consider the following code chunk.

```{r roll_die, echo=TRUE}
success = 0  # storage vector for whether a roll is a 2 or not
nrolls = 1000 # number of rolls of the die in the simulation experiment
for(i in 1:nrolls){
  roll = sample(1:6, 1) # roll the six-sided die once
  # Two ways to determine if the roll is a 2
  # In each case, we want to store the success result in 
  # the ith element of the vector `success'.
  # If roll is a 2, store value of '1' in the success vector; 
  # otherwise store a value of zero
  # 1) If-then statement: an if-else syntax for the Boolean expression
  if(roll==2){success[i]=1}else{success[i]=0}
  # 2) Straight Boolean expression
  #success[i] = (roll == 2) 
  }
mean(success) # proportion of 1s (successes) in the simulation experiment
```

Notice that we embed a single simulation experiment (roll a die) inside a for-loop to repeat the experiment many times (here 1000 die rolls).  After each experiment, we store an indicator of success of the experiment in a vector.  So here, the vector `success` is 1000-dimensional and has a value of 1 if a roll is a 2 and a value of 0 if the roll is not a two. 
Try running this code for 10 die rolls, and output the vector `success`.  Notice that it is a vector of 0s and 1s.  If we take an average of the success vector, we are computing the proportion of simulation experiments where a success occurred.  So in this example, an empirical probability that a 2 is rolled on a six-sided die.

In each of the following two tasks, you will repeat a simulation experiment by embedding code for a single simulation within a for-loop.  Each time, you need to store a success.  In the first task we will use an if-else statement.  In the second task we will use a Boolean expression.


## Task 1: Simulating the birthday problem

The birthday problem considers the probability that two people in a group of a given size have the same birth date.  We will assume a 365 day year (no leap year birthdays).

### Code set-up

Dobrow 2.28 provides useful R code for simulating the birthday problem.  Imagine we want to obtain an empirical estimate of the probability that two people in a class of a given size will have the same birth date.  The code

`trial = sample(1:365, numstudents, replace=TRUE)`

simulates birthdays from a group of `numstudents` students.  So you can assign `numstudents` or just replace `numstudents` with the number of students in the class of interest.

If we store the list of birthdays in the variable `trial`, the code

`2 %in% table(trial)`

will create a frequency table of birthdays and then determine if there is a match (2 birthdays the same).  We can use this code in an if-else statement to record whether a class has at least one pair of students with the same birth date.
We then can embed the code within a for-loop to repeat the experiment, store successes in a vector, and then take the average number of successes (a birthday match) across the repeated tasks.

### The problems

* Simulate the birthday problem to obtain an empirical estimate of the probability that two people in a class of 23 will have the same birth date.  In particular, simulate birthdays for 1000 classes (`for(i in 1:1000){...}`) each of size 23 and compute the proportion of these classes in which at least one pair of students has the same birth date.

##### Recall that the true probability is `1-prod(seq(343,365))/(365)^23` which is approximately 50%.

* Using your simulation code, estimate the number of students needed in the class so that the probability of a match is 95%.  (You may do this by trial and error.)

* Using your simulation code, find the approximate probability that three people have the same birthday in a class of 50 students.

```{r sim_birthdays, echo=TRUE}
BirthSim = function(n, m, c)
{
  x = replicate(n, m %in% table(sample(1:365, c, replace = TRUE)))
  return(mean(x))
}
# n trials, m students with same birthday, c students per class.

```

Place your answers to the three items below here:

* *The probability that 2 people in a class of 23 share a birthday is approximately: `r BirthSim(10000, 2, 23)`*

* *The probability that 2 people in a class of 48 share a birthday is approximately: `r BirthSim(10000, 2, 48)`*

* *The probability that 3 people in a class of 50 share a birthday is approximately: `r BirthSim(10000, 3, 50)`*

## Task 2: Random permutations

### Code set-up
A random permutation is a random shuffling of a set of integers $1, \ldots, n$. The following code chunk presents code for a random permutation algorithm from Dobrow Example 2.14. The code chunk outputs the original list and then the random permutation of that list of $n$ integers (in the code $n = 12$).  Try running it yourself and note the shuffling or permutation of the $n$ integers.

The algorithm sequentially moves down the list.  At each position, say $i$, it swaps the number in slot $i$ with a randomly chosen element from positions `i:n` (so later in the list or no swap at all).  
```{r rand_perm, echo=TRUE}
# Simulating random permutations
# Code from Example 2.14 of Dobrow (2014)
	n = 12  # permutation of size n
	perm = 1:n # store list of n integers
	# sequentially move down the list
	for(card in 1:(n-1)){
		x = sample(card:n,1) # randomly chose a position for swapping
		old = perm[card] # store the orginal number in position i
		perm[card] = perm[x] # replace the number in position i with the number in the randomly chosen position x
		perm[x] = old # complete the swap by placing the original number in position i in position x 
	}
1:n  # original list
perm # randomly permuted list
```

In this task, we are going to shuffle a deck of cards and determine the probability that the top (first element in permutation) and bottom (last element in permutation) card are the same.  The code

`(floor((perm[1]-1)/13) == floor((perm[52]-1)/13))`

checks if the first and last cards from the random permutation are the same suit.  After performing the random permutation, you can store this value in a success vector (Boolean expression), and then wrap the code in a for-loop to repeat the simulation.  The proportion of experiments reporting a 1 in the success vector is an estimate of the probability that the top and bottom cards of a shuffle are the same.

### The problems
Based on Dobrow problem 2.32.  Revise the code chunk above to shuffle a standard deck of cards (52 cards).  Simulate the probability that in a randomly shuffled deck, the top and bottom cards are the same suit.

```{r shuffle_deck, echo=TRUE}
FILO = function(n)
{
  shuffle = matrix(replicate(n, sample(1:52, 52, replace = FALSE)), nrow=52, byrow=FALSE)
  FLmatch = floor((shuffle[1,1:1000]-1)/13) == floor((shuffle[52,1:1000]-1)/13)
  return(mean(FLmatch))
}
```

### Questions:

* Present the empirical probability (based on your simulation experiment) that in a randomly shuffled deck the top and bottom cards are the same suit.

*The probability that the first and last card of a shuffled deck are the same suit is approximately `r FILO(10000)`*

* Briefly explain how the code `(floor((perm[1]-1)/13) == floor((perm[52]-1)/13))` checks if the first and last cards from a random shuffling (permutation) are the same suit.

*If the cards in each suit are represented by 13 consecutive numbers (1-13, 14-26, 27-39, 40-52) and represent each individual with the numbers 0, 1, 2, or 3; then subtracting 1 from the index of each card in the deck and rounding down the division by 13 will collapse each set of indices down to 0, 1, 2, or 3. If 2 different card indices fall to the same number then we may conclude they are the same suit.*

* Based on our discussion of counting and probability, present an intuitive explanation of why the probability follows the value given by your simulation.  (I.e., I am not asking you to mathematically find the probability.  Just briefly explain why the value you obtained is what you may expect it to be based on the probability material covered to this point.)

*The probability is approximately 25%, which makes sense when you consider that there are $4^2$ ways to choose 2 different suits for the first and last cards, and 4 ways to choose one suit. So, intuitively, the probability should be about $\frac{4}{4^2} = \frac{1}{4} = 25$%.*

